---
title: "Lab 6: Statistical tests: Mann-Whitney U, Wilcox tests, Linear Regression"
author: "Sophia Leiker"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```


## Rank based text example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test).

H0: Equal ranks (or medians) (null hypothesis)
HA: Ranks are not equal (or, medians are not equal)

### Create samples 

'set.seed()' to create "pseudorandom, but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributions using 'hist()'

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the argument 'paired = "TRUE"' within the t.test() function

```

This p-value means that IF our samples are drawn from populations with the same mean (i.e., if the null hypothesis is true), there is a `r round(ttest$p.value,2)` probability of finding sample means *at least as different* as mine by random chance. 

[A more comprehennsive explaination of difference between groups... actuall means, data spread, effect size...] there is no significant difference (t(df) = t-statistic value, p = )

### Mann Whitney U Rank Based test

We use the 'wilcox.test()' function to perform a two sample-rank based test

```{r}

mwu <- wilcox.test(gp_1, gp_2)

```


This p-value means that IF our samples are drawn from populations with the same median (i.e., if the null hypothesis is true), there is a `r round(mwu$p.value,2) * 100` % probability of finding sample medians *at least as different* as mine by random chance. 

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value, 2) `)

## Simple linear regression

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense. 

Use the 'penguins' data from {palmerpenguins}, exploring flipper length (x), versus body mass (y).

```{r}
# Use view(penguins) to look at it
# Make an exploratory scatterplot of penguin flipper length (x-axis) versus body mass (y-axis).

penguin_scatter <- ggplot(data = penguins, aes(x= flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = lm)

penguin_scatter

```


Sophia Notes:
body_mass (g) = B1(flipper_length(mm)) + B2
B1 --> Slope, B2 --> Intercept. Linear regression minimizes the residuals (i think between the true results and model estimates)

### Find the model estimates (coefficients)

Use the `lm()` function for a linear model

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

```

Sophia Notes: Intercept is -5780.83 and this is in g, slope of flipper_length_mm 49.69 is in units of g/mm. This is because the grams on the right means there has to be grams on the left. This means that on average if the penguins gain 1 gram in body mass, then their flipper length will increase by 49.69mm. The expected average mass of penguins with flipper length of 0 is the intercept, -5780.83 g. 


Model coefficients: `r penguin_lm$coefficients`
Flipper length coefficient: `r penguin_lm$coefficients [2]` 

#INDEXING: using square brackets here to select only the second element of the model coefficient lists 

### Use `broom::tidy()' to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```

### Use broom::glance() to get output of model fit and statistical significance


```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance #this provides us with a ton of stats
```

### Use equatiomatic::extract_eq() to return model in LaTeX

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE) #this second part on use_coefs means you want to automatically include the values in the expression
```

### Use plot() function to automatically produce diagnostic plots

```{r}
plot(penguin_lm)
```


